Comparing ChatGPT-4 to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6939 Deaths in Sierra Leone from 2019-2022

### 2024-02-09 manuscript/manuscript.tex
Add perf overall by age group
Add model specs and programming versions; update methods diagram

Initially, 11,920 records from the HEAL-SL study \cite{njalauniversityHealthySierraLeone2023,carshon-marshChildMaternalAdult2022} were collected from dual-coded EVA, where each record was randomly coded by two different physicians that assigned CODs as International Classification of Diseases Revision 10 (ICD-10) codes \cite{worldhealthorganizationICD10InternationalStatistical2011}. Physicians were able to assign CODs for 11,820 of the 11,920 records, where 100 of these records could not be assigned a COD due to missing or inadequate information (e.g. low quality narrative, data loss). To determine if two codes were in physician agreement per record, codes were compared for similarity using Central Medical Evaluation Agreement 10 (CMEA-10) codes, which groups a range of similar ICD-10 codes together indicating codes in agreement \cite{aleksandrowiczPerformanceCriteriaVerbal2014} (see Additional File 2). When codes were not in agreement, a record enters the reconciliation phase, where the two physicians were provided reasoning and initial codes from each other to: (1) keep their initial code (2) assign the other physician's code or (3) assign a new code. If codes were not in agreement after the reconciliation phase, a record enters the adjudication phase, where a third senior physician evaluates both physicians' reasoning and codes before and after reconciliation, and assigns a final code based on their evaluation. The 11,820 physician coded records were further filtered for records where both physicians agreed on the assigned codes (records that were not reconciled or adjudicated) resulting in 6942 physician agreed records. Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \cite{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}. The 6942 records were converted into CGHR-10 codes (see Additional File 1) that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 years or older), child (28 days to 11 years), and neonatal (under 28 days) age groups.  After conversion, a final total of 6939 physician agreed records (3826 adult, 2636 child, and 477 neonatal) were used for modelling and performance evaluation, where three records were removed as their ICD-10 codes did not have a matching CGHR-10 code.

@@

\subsubsection{Model Parameters}\label{methods-models-params}

The parameter \verb|top_p| for ChatGPT-3.5 and ChatGPT-4, representing the sampling temperature, was set to 0.3\footnote{We tested for top\_p values between 0.1 to 0.9 at intervals of 0.1 on five records that did not return outputs from ChatGPT-4 until all five records returned output. At top\_p=0.3, all five records produced output.} to consider only the top 30\% probability mass from the token results \cite{openaiOpenAIPlatformAPI2024}. Higher values closer to 1 may make the results more random, while lower values closer to 0 make the results more focused. The \verb|Nsim| (number of iterations to run) parameter \cite{liInSilicoVAProbabilisticVerbal2022} for InSilicoVA was set to 9500\footnote{The default value of Nsim=10000 for InSilicoVA ran until 9500 iterations before it stopped due to errors, thus Nsim=9500 was used and ran successfully for all iterations.}, while the \verb|HIV| (level of prevalence of human immunodeficiency virus) and \verb|Malaria| (level of prevalence of Malaria) parameters \cite{thomasInterVA5ReplicateAnalyse2021} for InterVA-5 were both set to \verb|'h'| (high) reflecting HIV and Malaria disease assumptions in Sierra Leone \cite{yendewaHIVAIDSSierra2018,walkerMalariaMorbidityMortality2015}.

@@

>>>>

Initially, 11,920 records\footnote{Available at openmortality.org/dataset/heal-sl} from the HEAL-SL study \cite{njalauniversityHealthySierraLeone2023,carshon-marshChildMaternalAdult2022} were collected from dual-coded EVA, where each record was randomly coded by two different physicians that assigned CODs as International Classification of Diseases Revision 10 (ICD-10) codes \cite{worldhealthorganizationICD10InternationalStatistical2011}. Physicians were able to assign CODs for 11,820 of the 11,920 records, where 100 of these records could not be assigned a COD due to missing or inadequate information (e.g. low quality narrative, data loss). To determine if two codes were in physician agreement per record, codes were compared for similarity using Central Medical Evaluation Agreement 10 (CMEA-10) codes, which groups a range of similar ICD-10 codes together indicating codes in agreement \cite{aleksandrowiczPerformanceCriteriaVerbal2014} (see Additional File 2). When codes were not in agreement, a record enters the reconciliation phase, where the two physicians were provided reasoning and initial codes from each other to: (1) keep their initial code (2) assign the other physician's code or (3) assign a new code. If codes were not in agreement after the reconciliation phase, a record enters the adjudication phase, where a third senior physician evaluates both physicians' reasoning and codes before and after reconciliation, and assigns a final code based on their evaluation. The 11,820 physician coded records were further filtered for records where both physicians agreed on the assigned codes (records that were not reconciled or adjudicated) resulting in 6942 physician agreed records. Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \cite{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}. The 6942 records were converted into CGHR-10 codes (see Additional File 1) that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 years or older), child (28 days to 11 years), and neonatal (under 28 days) age groups.  After conversion, a final total of 6939 physician agreed records (3826 adult, 2636 child, and 477 neonatal) were used for modelling and performance evaluation, where three records were removed as their ICD-10 codes did not have a matching CGHR-10 code.

@@

\subsubsection{Models and Parameters}\label{methods-models-params}

The ChatGPT-3.5 and ChatGPT-4 Application Programming Interface (API) was accessed using Java and used to assign CODs for each record. ChatGPT-3.5 used the \verb|gpt-3.5-turbo| model, while ChatGPT-4 used the \verb|gpt-4-0613| mode. The parameter \verb|top_p| for ChatGPT-3.5 and ChatGPT-4, representing the sampling temperature, was set to 0.3\footnote{We tested for top\_p values between 0.1 to 0.9 at intervals of 0.1 on five records that did not return outputs from ChatGPT-4 until all five records returned output. At top\_p=0.3, all five records produced output.} to consider only the top 30\% probability mass from the token results \cite{openaiOpenAIPlatformAPI2024}. Higher values closer to 1 may make the results more random, while lower values closer to 0 make the results more focused.

The \verb|openVA| R package was used to run InterVA-5 and InSilicoVA models to assign CODs for each record in R version 4.3.1. The \verb|openVA| package version 1.1.1 used dependent packages \verb|InterVA5| version 1.1.3 and \verb|InSilicoVA| version 1.4.0. The \verb|Nsim| (number of iterations to run) parameter \cite{liInSilicoVAProbabilisticVerbal2022} for InSilicoVA was set to 9500\footnote{The default value of Nsim=10000 for InSilicoVA ran until 9500 iterations before it stopped due to errors, thus Nsim=9500 was used and ran successfully for all iterations.}, while the \verb|HIV| (level of prevalence of human immunodeficiency virus) and \verb|Malaria| (level of prevalence of Malaria) parameters \cite{thomasInterVA5ReplicateAnalyse2021} for InterVA-5 were both set to \verb|'h'| (high) reflecting HIV and Malaria disease assumptions in Sierra Leone \cite{yendewaHIVAIDSSierra2018,walkerMalariaMorbidityMortality2015}.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.7\textwidth]{fig-perf-agegroup.pdf}
\caption{Model performance by age group. ChatGPT-3.5, ChatGPT-4, InterVA-5, and InSilicoVA model performance for assigning CODs in the adult (12 to 69 Years), child (28 days to 11 years), and neonatal (under 28 days) age groups. CSMF indicates CSMF accuracy.}\label{fig-perf-agegroup}
\end{figure}
